---
title: 'Regression Analysis: Weather Forecasts'
author: "Claudia Chu"
date: "2022-08-02"
categories: R
tags:
- Regression
- Data Visualisation
- R Markdown
---
In this project, I will be running regression analyses on the Weather in Szeged dataset (2006-2016) from Kaggle. 

# Setup
```{r Load packages, message = FALSE, warning = FALSE}
library(tidyverse) # data wrangling
library(broom) # data wrangling
library(lubridate) # data wrangling
library(ggplot2) # graph plotting
library(ggloop) # graph plotting
library(corrplot) # heat map
library(car) # regression
```

```{r Load data file, message = FALSE, warning = FALSE}
# Load file
dataset <- read.csv("weatherHistory.csv", header = TRUE, na.strings = c(""," ","NA")) 

# Format the variable names
colnames(dataset) <- gsub(x = colnames(dataset), pattern = "\\.", replacement = "_")
```

First, let's take a look at the data. 
```{r}
summary(dataset)
```
There is no variability in the variable "Loud Cover", so it will be excluded in subsequent analysis. Similarly, "Formatted Date", "Summary, "Daily Summary", and "Precipitation Type" are not of my interest in this linear regression, so they will be excluded as well. "Temperature" and "Apparent Temperature" are of similar construct, so the latter is removed from the analysis as well.

```{r}
# Select the variables of interest
dataset <- dataset %>% 
           select(-c("Formatted_Date", "Precip_Type", "Apparent_Temperature__C_", "Loud_Cover", "Daily_Summary", "Summary"))

# Rename the variables for better clarity
dataset <- dataset %>% 
           rename(Temperature = Temperature__C_,
                  Wind_Speed = Wind_Speed__km_h_,
                  Wind_Bearing = Wind_Bearing__degrees_,
                  Visibility = Visibility__km_,
                  Pressure = Pressure__millibars_)
```
Now that I have selected the variables from the data set, I can start screening the data. 

First, let's check for missing data.
```{r}
sum(is.na(dataset))
```
The data set does not consist any missing values. 

Next, let's look at outliers for the variables at interval and ratio levels. This can be done using box plots. 
```{r}
ggplot2::theme_set(ggplot2::theme_light())  ## Globally sets the ggplot2 theme

boxplots <- ggloop(data = dataset, 
                mappings = aes_loop(x = "", y = Temperature:Pressure))

boxplots[1:6] <- boxplots[1:6] %L+% geom_boxplot(colour = "grey35", fill = "steelblue")
boxplots[1:6]
```

Removing outliers is not ideal, so outliers in this data set will be replaced with the 5th and 95th percentiles respectively. 
```{r}
replace_quantile <- function(x){
    quantiles <- quantile( x, c(.05, .95 ) )
    x[ x < quantiles[1] ] <- quantiles[1]
    x[ x > quantiles[2] ] <- quantiles[2]
    x
}

dataset$Temperature <- replace_quantile(dataset$Temperature)
dataset$Humidity <- replace_quantile(dataset$Humidity)
dataset$Pressure <- replace_quantile(dataset$Pressure)
dataset$Wind_Bearing <- replace_quantile(dataset$Wind_Bearing)
```

# Correlation Heatmap
To visualise the relationships of the indicators, I produced a correlation heat map. 
```{r}
corrplot(cor(dataset), method = 'shade', order = 'FPC', diag = FALSE, type = "lower")
```
From the graph, it is shown that temperature and humidity had a strong negative correlation. 

# Regression
Next, I will run the multiple linear regression analyses on the data set, with the aim to predict temperature.
```{r}
Model <- lm(Temperature ~ Pressure + Humidity + Visibility + Wind_Bearing + Wind_Speed, dataset)
summary(Model)
```
Results showed that pressure, humidity, visibility, wind bearing and wind speed all significantly predicted temperature. The model explained 54% of the variances. 
