---
title: 'Linear Regression Analysis'
author: "Claudia Chu"
date: "2022-08-02"
categories: R
tags:
- Regression
- Data Visualisation
- R Markdown
---
In this project, I will be examining some of the predictors of countries' total adult literacy rates using the world bank data. Multiple linear regression analyses will be conducted. The indicators of interest are: a) government expenditure on education, b) trained teachers in the primary education sector, c) pupil-teacher ratio in the primary education sector, d) fertility rate, e)  GDP per capita, f) Gini index, g) research and development expenditure, h) school enrollment percentage in the primary education sector

# Setup
```{r Load packages, message = FALSE, warning = FALSE}
library(tidyverse) # data wrangling
library(broom) # data wrangling
library(ggplot2) # graph plotting
library(car) # regression
library(lmtest) # regression
library(naniar) # missing data analysis
library(mice) # multiple imputation
```

```{r Load data file, message = FALSE, warning = FALSE}
# Load file
dataset <- read.csv("Literacy.csv", header = TRUE, na.strings = c(""," ","NA")) 

# Format the variable names
colnames(dataset) <- gsub(x = colnames(dataset), pattern = "\\.", replacement = "_")
```

# Data Cleaning
Before running the regression, the data has to be screened and cleaned. First of all, missing data has to be addressed. Countries with missing data on the outcome measure, i.e., the total adult literacy rates, will be excluded from subsequent analysis. Such decision was made to avoid creating any bias in drawing conclusions from the results. It could be argued that the data set will be less representative of the world population, but such trade-off is unavoidable and future improvements could be made by collecting more data. 

A total of 32 countries were left after applying the exclusion criteria. 
```{r}
dataset <- dataset %>% filter(!is.na(Literacy_Rate))


gg_miss_upset(dataset, nsets = n_var_miss(dataset))
```

